---
title: "USEFUL FUNCTIONS"
output:
  rmdformats::downcute:
    downcute_theme: "chaos"
---

```{r include=FALSE}
#above makes this code chunk not show up in output document, but RUNS the code below. 
#below sets to print all other code blocks, but not run (evaluate/eval) any of them. 
knitr::opts_chunk$set(eval = FALSE)
```

hit control + shift + C to comment-out entire highlighted section of code (and un-comment it)

# data import/export
```{r}
saveRDS(bad_dates, "bad_dates.rds")  
```







standard csv read-in. Automatically imports at tibble.

```{r}
read_csv("filename.csv")%>%
  select(-1)
```

standard csv write with no rownames.

```{r}
write_csv("filename.csv")
```



Save R objects that take a long time to read/write to other file types,
Or objects that don't work conveniently for CSV files. 
```{r}
saveRDS(my_precious, here("results", "my_precious.rds"))
my_precious <- readRDS(here("results", "my_precious.rds"))
```


















# df manipulation

move row to column names, uses janitor package. takes number for row number.

clean_names, also from janitor. makes titles R-friendly.

```{r}
df%>%
  row_to_names(1)%>%clean_names()
```


remove column that is ALL NAs. 
```{r}
 select(where(function(x) any(!is.na(x))))
```





1 way to get dates from dates.

```{r}
mutate(col = as.Date(col, format = "%m/%d/%Y"))
```

in new column, delete everything before the space.
Or delete special characters. Use 2 backslashes. 

```{r}

  mutate(injection_date = gsub(" .*", "", injection_date))
mutate(necropsy_file_path = gsub("\\~\\$", "", necropsy_file_path))
```

mutate if data in x column if y column contains a string/pattern.

use grepl here to turn the search into a logical vector of same length as search area, so that if_else can work.

Tells R to leave any data in x column alone, if search string isn't found in y column.

```{r}
 mutate(variable.x  = if_else(grepl("string.search", variable.y), "string.result", variable.x))
```

mutate conditional values

```{r}
mutate(gender = if_else(gender == "Male", "M",
                if_else(gender == "Female", "F",
                gender)))
```

mutate_at to mutate multiple columns using the same function. can also use helper functions to select columns.

here mdy is the function, uses vars to select the variables, in this case by column index.

```{r}
mutate_at(vars(13:17,23), mdy)
```

or

```{r}
cohorts2.1%>%
  mutate_at(vars(name1:name5), ~gsub("pwt", "+/+", .x))
```






extract filename from path using gsub function.

```{r}
mutate(nec_filename = gsub(".*\\/", "", necropsy_file_path))
```

path sanitize to remove all "illegal" characters from potential filename. 
```{r}
library(fs)
path_sanitize()
```





getting means across rows, leaving NAs out of the calculations. 
```{r}
#make vars numeric if not already.
mutate_at(c("diff1", "diff2", "diff3"), ~as.numeric(.x))%>%
  #code allows calcuation to ignore NAs. 
  mutate(diff = round(rowMeans(across(c("diff1", "diff2", "diff3")), na.rm=TRUE), 1))
```









# plotting

re-order x-axis:

create vector with desired order of x-axis. Order axis by setting "level" vector within factor() function in x aesthetic.

```{r}
x_order <- c("pre_tumorigenic", "glioma","spinal_cord", "head", "nerve")
             
ggplot(aes(x = factor(tumor_location, level = x_order), 
            y = this))
```

#allows border

```{r}
shape=1 
```

#can be filled.

```{r}
shape=16 
```

position dodge

```{r}
position = position_dodge(width = 1)
```

# file manipulation

copy files from and to.

```{r}
old<-("directory")
new<-("directory")
file.copy(from=old, to=new, 
          overwrite = TRUE, 
          recursive = TRUE, 
          copy.mode = TRUE)
```

list files
#if using a subdir of wd, must use "."

```{r}
list.files(path = "directory", 
           full.names = TRUE, 
           recursive = TRUE, 
           ignore.case = TRUE)%>%
  as_tibble()



list.files("./necropsy/",pattern = "*.csv") 


#use "../" to go up one from current wd. 
list.files("../slides/file.csv") 

```

rename files

```{r}
file.rename(filenames$new, filenames$old)
```


```{r}
 #code to turn name of file into working hyperlink (in excel)
   mutate(link = paste0("=HYPERLINK(", "\"", 
                      #full path to file
                       path,
                      #add  [","]
                      "\",\"", 
                      #name of link
                       filename,
                      #add [")]
                      "\")"))
```

